---
title: "Introduction SQL in RStudio, part 2, teacher version"
date: June 2019
author: Antoinette Littel
number_sections: true
code_folding: hide
output: html_notebook

---
In part 2 were are focussing on the aggregate functions (COUNT, MAX, MIN, AVG, SUM) in combination with the GROUP BY. 


<h2> Connecting to the database </h2>

The first chunk in this Notebook will run automatically when you make the SQL statements. It loads the libraries you need and connects to the database 'CapeCodd.db' that should be in your folder. 

```{r setup, results='hide'}
library(tidyverse)
library(readxl)
library(DBI)

#enable other connections
# dbDisconnect(con)

# connect to CapeCodd
con <- DBI::dbConnect(RSQLite::SQLite(), "CapeCodd.db")
#

#
# use the database connection for all SQL chunks in this notebook and hide the code by default
knitr::opts_chunk$set(connection = "con", echo=FALSE)

```



<h2>Database design </h2>

Below is the design of the database. You should have this picture in the images folder of your workdirectory. You can open the picture next to your Notebook, so that you can see the design while creating your queries.

<img src = "images/CapeCodd.png" />


<h2> Assignments </h2>

Make a SQL query for the following questions. 

A. What is the total price per order (use ORDER_ITEM)?
```{sql, connection=con}
SELECT * FROM ORDER_ITEM

```


<br>


B. What is the total price per order, for orders that have total quantity of items more than 2?
```{sql, connection=con}
SELECT ordernumber, SUM(ExtendedPrice) FROM ORDER_ITEM
GROUP BY(ordernumber)
HAVING sum(quantity) > 2
```
<br>

C. What is the total price per order, for orders with more than 2 different items?
```{sql, connection=con}
SELECT ordernumber, SUM(ExtendedPrice) FROM ORDER_ITEM
GROUP BY(ordernumber)
HAVING sum(quantity) > 2
```
<br>

E. For which SKU's the <b>average</b> QuantityOnHand is greater than the <b>highest</b> QuantityOnOrder for any Warehouse?
```{sql, connection=con}


```
<br>

F. For which Warehouses the total QuantityOnHand is greater than the total QuantityOnOrder?
```{sql, connection=con}


```
<br>

G. For which Warehouses the total QuantityOnHand for Tents is greater than half the total QuantityOnOrder?
```{sql, connection=con}



```
<br>

H.	Give the total quantity of all sku items in all the warehouses. Name the columns as sku, sku_description and QOH?
```{sql, connection=con}


```
<br>

<h2>Extra practice</h2>
<h3> GROUP BY, multiple columns</h3>

E1. How many items(SKU's) where ordered per price?
```{sql, connection=con}


```
<br>

E2.	How many different items(different SKU codes) are in the inventory per warehouse?
```{sql, connection=con}


```
<br>

E3.	How many different items(different SKU codes) are in the inventory for Warehouse 300?
```{sql, connection=con}


```
<br>
E4.	How many different items(different SKU codes) are in the inventory for Warehouse 300? But now, count only the items where the QuantityOnHand is higher than 50.
```{sql, connection=con}

```
<br>

E5.	How many items are there in total (use QuantityOnHand) in the inventory per warehouse?
```{sql, connection=con}


```

<h2> Challenging assignments</h2>



C1. Visualize the QuantityOnHand per SKU. Make two chunks: a sql chunk and a R chunk for <code> ggplot </code>

```{sql, connection = con, output.var = "df_result", echo=FALSE}



```
Remember to use the <code>{r}</code> tag in your chunk to switch to regular R again. 

```{r, echo=FALSE}

```
